---
title: "Cherry Blossoms, Nepali Study"
author: "Michael Cowan"
format: 
  html:
    embed-resources: true
    df-print: paged
editor: visual
---

```{r}
library(tidyverse)
```

## 1. The Cherry Blossom Run

a\. This is a .csv file, so it will be simple to read in.

```{r}
cherry_blossoms <- read.csv("http://ritsokiguess.site/datafiles/run12.csv")
cherry_blossoms
```

b\. Below is a histogram of the time taken to run the race, including all the runners together. It would be best to use the median as a measure of centre than the mean due to it not being influenced by outliers or skewed data. Specifically, in the context of a race, if there are a few runners with exceptionally fast, or slow, running times, using the mean as a measure of central tendency could give a distorted picture of the measure of centre (the typical running time of the participants in the race).

```{r}
ggplot(cherry_blossoms, aes(x = time)) + geom_histogram(bins = 10)
```

c\. Here is a histogram for time for males and female runners. The shapes are somewhat similar, but it is evident that the male distribution is left-skewed, while the female distribution appears to be right skewed. Thus, there seem to be more faster male runners than female runners

```{r}
ggplot(cherry_blossoms, aes(x = time)) + geom_histogram(bins = 10) + facet_wrap(~gender)
```

d\. Below are the summaries of the median and interquartile ranges of times for all runners together.

```{r}
cherry_blossoms %>% summarize(median_time = median(time), IQR = IQR(time))
```

e\. Here are the number of runners, the median, and the interquartile range of times for male and female runners separately. Specifically, there were approximately 2500 more female participants in the race. Females took on average, approximately 10 minutes longer than male runners. The IQR for male runners was 20.22 minutes -- slightly higher than the 18.82 minutes for female runners -- indicative of more variability in running times among the middle 50% of male runners compared to the female runners.

```{r}
cherry_blossoms %>% group_by(gender) %>% summarize (number_of_runners = n(), median_time = median(time), IQR = IQR(time))
```

## 2. Nepali child health study

a\. Once again, we're dealing with a .csv file. I've read in the data and displayed it .

```{r}
child_study <- read.csv("http://ritsokiguess.site/datafiles/nepali.csv")
child_study
```

b\. Let's display the weights and heights for all the children with the select function.

```{r}
child_study %>% select(wt, ht)
```

c\. Now we'll display the weights and heights and mother's literacy for all children, without naming any of the columns. With a quick glance at the columns, only the columns representing weight, height, and literacy end with the letter "t", so the select "ends_with" function can achieve this goal.

```{r}
child_study %>% select(ends_with("t"))
```

d\. Below, I'll display all the data for which the child is female by filtering by sex (using the == logical condition).

```{r}
child_study %>% filter(sex == "female")
```

e\. Now I'll display the child's age and weight for the children born of only illiterate mothers, selecting age and weight.

```{r}
child_study %>% filter(lit == "illiterate") %>% select(age, wt)
```

f\. Now I'll make a graph comparing the weights of the children born to illiterate vs. literature mothers, allowing for the fact that the children were measured at different ages. I'll put a linear regression line through the data with geom_smooth to better visualize the relationships.

```{r}
# When trying to run geom_smooth R was giving me errors about the missing data, so the drop_na function came in handy here.
clean_child_study <- child_study %>% drop_na(age, wt, lit)
clean_child_study %>% ggplot(aes(x = age, y = wt, color = lit)) + geom_point() + geom_smooth(method = "lm")
```

Assuming this is a linear relationship between weight and age whether the mother is literate or not, we can conclude we can conclude that children of literate mothers are seemingly higher weight, on average, than those from illiterate mothers. In addition, there seems to be less dispersion in the weights of children from literate mothers than those born from illiterate mothers. Perhaps this is a reflection of another related variable -- such as socioeconomic status -- connecting parental literacy and child weight, but further analysis and inputing controls into the model would be necessary to explore this relationship.
