---
title: "STAC32 - Assignment 7"
author: "Michael Cowan - 997887032"
format: 
  html:
    embed-resources: true
    df-print: paged
editor: visual
---

```{r}
library(tidyverse)
library(devtools)
install_github("nxskok/smmr")
library(smmr)
```

# **1. Hurricanes**

a\. I'll read in the data with read_csv.

```{r}
hurricanes_data <- read_csv("http://ritsokiguess.site/datafiles/hurricanes.csv")
hurricanes_data
```

b\. I'll start with a boxplot of the hurricanes by time period.

```{r}
ggplot(hurricanes_data, aes(x = period, y = Hurricanes)) + geom_boxplot() + 
  labs(x = "Time Period", y = "Number of Hurricanes")
```

There appear to be differences in the median hurricanes per time period between the four groups. That said, we'll have to do an ANOVA to see if at least one of the group differences is statistically significant (followed by multiple comparisons to determine which groups specifically differ).

Given we're going to do an ANOVA, we'll have a quick glance at a normal quantile plot to see the normality of the data (below). I'll accept that the majority of the plot points fall on the reference line and proceed with the ANOVA.

```{r}
ggplot(hurricanes_data, aes(sample = Hurricanes))+ stat_qq() + stat_qq_line() + facet_wrap(~period)
```

c\. Given that we've accepted the data to be visually "normal enough" (from the quantile plot), we're justified in running an ANOVA to compared the number of hurricanes in the four time periods.

```{r}
hurricanes.aov <- aov(Hurricanes~period,data=hurricanes_data)
summary(hurricanes.aov)
```

The ANOVA meets the threshold for significance (p = 0.00519). This indicates that there are statistically significant differences in the average number of hurricanes across at least one of the different time periods (we reject the null hypothesis). Therefore, we'll proceed with multiple comparison -- specifically, Tukey's HSD to control for family-wise alpha (given we're conducting multiple tests, this is appropriate to control the error rate).

```{r}
TukeyHSD(hurricanes.aov)
```

As we can see from the results, the only significant difference in hurricanes per time period are between the 1990-2014 period and the 1914-1939 period, p = 0.0030575 (reject the null hypothesis). Specifically, the test indicates there is a significant mean difference -- an average of \~2.5 more hurricanes -- between 1990-2014 compared to the 1914-1939 period. The significant increase in hurricane frequency between the earliest and latest time periods, contrasted with the lack of significant changes in the intervening years, could be indicative of the notable increase in global temperatures witnessed in the late 20th and early 21st century -- which might correlate with increased hurricane activity.

# **2. Nitrofen**

a\. We'll read in the data with read_csv.

```{r}
herbicide <- read_csv("http://ritsokiguess.site/datafiles/nitrofen.csv")
herbicide
```

b\. We'll have to use pivot_longer to arrange the data so the numbers of offspring are all in one column, labelled by which brood they are from.

```{r}
herbicide_longer <- herbicide %>%
  pivot_longer(cols = starts_with("brood"), names_to = "brood", values_to = "offspring")
herbicide_longer
```

c\. Now we'll use a spaghetti plot to plot the number of offspring against the brood number, joining the points for the same animal with lines, and distinguishing Nitrofen Concentration by colour.

```{r}
ggplot(herbicide_longer, aes(x = brood, y = offspring, group = animal, color = factor(conc))) + geom_line() + geom_point() + labs(title = "Offspring Count per Brood by Nitrofen Concentration", x = "Brood Number", y = "Number of Offspring", color = "Nitrofen Concentration (µg/L)")
                                                                                                                                  
```

d\. Ultimately, the graph indicates that the reproductive success of Ceriodaphnia dubia decreases with increased Nitrofen concentration levels (higher concentrations of Nitrofen = less reproductive success). Specifically, those Ceriodaphnia dubia allocated to higher concentrations of Nitrofen appeared to reproduce less on average, when compared to those allocated to lower Nitrofen concentrations. For example, this relationship is most apparent when comparing the 310 micrograms/L concentration (the pink line) who produced zero offspring by Brood 3, compared to those in the 0 micrograms/L concentration condition, who produced at a minimum 10 offspring by Brood 3.
