---
title: "Choices by Demography, NBA TV"
author: "Michael Cowan"
format: 
  html:
    embed-resources: true
    df-print: paged
editor: visual
---

```{r}
library(tidyverse)
library(MASS) # issue with "select", that conflicts with our reg. select.
library(marginaleffects)
library(broom)
library(nnet) # for multinom
library(conflicted) # To deal with MASS, we use "conflicted" package.
conflict_prefer("select", "dplyr") # whenever you run into "this", use the one from tidyverse ("dplyr").
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("summarize", "dplyr")
```

## 1. Choice-box

**a.** We'll use read_csv here:

```{r}
choice_box1 <- read_csv("http://ritsokiguess.site/datafiles/Boxes.csv")
choice_box1
```

**b.** Multinom from the nnet package is used for **multinomial logistic regression, which is appropriate for relationships between a categorical dependent variable** **with more than two categories** and one or more IVs -- specifically, assuming the response categories are nominal (meaning they do not have a natural order/ranking to them).

**c.**

```{r}
choice_box1 %>% mutate(y = factor(y), age_squared = (age)^2) -> choice_box2
choice_box2
```

```{r}
choice_box.1 <- multinom(y ~ . , data = choice_box2)
```

**d.**

```{r}
choice_box.2 <- step(choice_box.1)
# In this case, drop1 would not work - due to a bug in multinom.
```

```{r}
choice_box.2
```

**e.**

R takes the values that are the most common values for each of the variables. Specifically, this is "boy" for the gender variable, "no" for the majority_first variable. In the case of age_squared (continuous), R takes the mean (or median) of the values.

-   Without age-squared assumes the effect is just straight up; age-squared allows it to bend.

```{r}
choice_box_frame <- datagrid(model = choice_box.2, age = c(5:13))
choice_box_frame
```

**f.**

Predictions: select the output columns you want to keep, and then pivot the group (predicted response category) wider:

```{r}
cbind(predictions(choice_box.2, newdata = choice_box_frame)) %>%
  select(group, age, gender, majority_first, estimate) %>%
  pivot_wider(names_from = group, values_from = estimate)
```

**g.**

```{r}
plot_predictions(model = choice_box.2, condition = c("age", "group"))
```

## 2. NBA Schedule

**a.**

```{r}
sched <- read_csv("http://ritsokiguess.site/datafiles/nba_sched.csv")
sched
```

**b.**

```{r}
sched %>% mutate(new_date = mdy(date), month = month(new_date), day = day(new_date), year = year(new_date), day_of_week = wday(new_date, label = TRUE)) -> sched_2
sched_2
```

```{r}
sched_2 %>%
  group_by(day_of_week) %>%
  summarise(number_of_games = n()) %>%
  arrange(desc(number_of_games)) -> sched_3
sched_3
```

**c.**

```{r}
NZ_games <- sched_2 %>% unite(date_time, new_date, start_eastern, sep = " ") %>% mutate(dt_EST = ymd_hm(date_time, tz = "America/Toronto")) %>% mutate(dt_auckland = with_tz(dt_EST, tz = "Pacific/Auckland"))
NZ_games
```

```{r}
games_after_4AUCK <- NZ_games %>%
  filter(hour(dt_auckland) >= 16)
games_after_4AUCK
```

There are 132 games (from 12/26/23 through 3/8/24) that the friend who lives in Auckland can watch. Primarily, these look like home games hosted on the West coast (primarily in California and Arizona), that traditionally have a later start time compared to East coast hosted games (ex. Toronto, Boston).
