---
title: "Cats, University Grades"
author: "Michael Cowan"
format: 
  html:
    embed-resources: true
    df-print: paged
editor: visual
---

```{r}
library(tidyverse)
library(MASS, exclude = "select") # issue with "select", that conflicts with our reg. select.
library(marginaleffects)
library(broom)
library(car) # for Levene's test
library(nnet) # for multinom
library(survival)
library(survminer)
library(MVTests)
library(lme4)
library(conflicted) # To deal with MASS, we use "conflicted" package.
conflict_prefer("select", "dplyr") # whenever you run into "this", use the one from tidyverse ("dplyr").
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("summarize", "dplyr")
```

## 1. Cats

**a.**

```{r}
cats <- read_csv("http://ritsokiguess.site/datafiles/cats.csv")
cats
```

**b.**

```{r}
cats_plot <- ggplot(cats, aes(x = Bwt, y = Hwt, colour = Sex)) +
  geom_point() + geom_smooth(method = "lm")
cats_plot
```

**c.**

```{r}
cats %>%
  group_by(Sex) %>%
  summarize(mean_Hwt = mean(Hwt), mean_Bwt = mean(Bwt))
```

```{r}
cats.1 <- lm(Hwt ~ Bwt * Sex, data = cats) 
anova(cats.1)
```

**d.** *Is the interaction term significant? What does your answer mean in the context of the data?*

The interaction term -- Bwt:Sex -- is statistically significant (p-value = 0.04722, \< 0.05), suggesting that a cat's body weight is different for different sexes (increase in heart weight with increasing body weight is not uniform for male and female cats).

**e.**

```{r}
cats_grid <- datagrid(Bwt = c(2.5, 3.5), Sex = c("M", "F"), model = cats.1)
cats_grid
```

```{r}
cbind(predictions(cats.1, newdata = cats_grid)) %>% 
  select(Sex, Bwt, estimate)
```

**f.** *Using your predictions, verify that the slopes for males and females are different.*

Visibly (below), the slopes for male and female cats are indeed different.

```{r}
plot_predictions(cats.1, condition = c("Bwt", "Sex"))
```

## 2. Sociology Grades

**a.**

```{r}
sociology <- read_csv("http://ritsokiguess.site/datafiles/socgrades.csv")
sociology
```

**b.**

Assumptions with BoxM (will refer to later):

```{r}
sociology %>% select(final, midterm1, midterm2) -> numeric_values
summary(BoxM(numeric_values, sociology$hssoc))
```

-   **Note**: The covariance matrices of the DV (the scores on the two midterms and the final) are not significantly different across the groups defined by the IV (hssoc) -- which also indicates that a MANOVA is suitable.

A MANOVA analysis is suitable given we are interested in student's subsequent performance scores on midterm1, midterm2, and final -- three quantitative dependent variables. Unlike an ANOVA, which assesses one DV, MANOVA lets us evaluate the effect of the IV (hssoc) on multiple DV outcomes.

Furthermore, MANOVA assumes performances across the three tests may be correlated. For our purposes, a MANOVA allows for understanding how previous exposure (hssoc) influences overall performance across the three assessments.

**c.**

```{r}
response <- with(sociology, cbind(midterm1, midterm2, final))
response
```

**d.**

```{r}
sociology.1 <- manova(response ~ hssoc, data = sociology)
summary(sociology.1)
```

**e.** *What do you conclude from your analysis, in the context of the data?*

Given the output, the p-value (0.03076) suggests statistical significance at 0.05 level. Thus, there is a statistically significant difference in the combined dependent variable (the scores) based on whether or not a student took high school sociology; taking hssoc does impact students' performance in the university sociology course.

**f.** *Make a graph that shows the effect of hssoc on each of the three exam scores. What does your graph say about the effect of hssoc on exam scores? (Hint: rearrange the data so that all the exam scores are in one column, and then use facets.)*

-   Code taken from: https://ritsokiguess.site/pasias2/manova.html#do-characteristics-of-urine-depend-on-obesity-1

```{r}
sociology %>%
  pivot_longer(col = c(midterm1, midterm2, final), names_to = "exam", values_to = "score") %>%
  ggplot(aes(x = hssoc, y = score, fill = hssoc)) + 
    geom_boxplot() +
    facet_wrap(~exam, scales = "free_y") +
    labs(x = "High School Sociology", y = "Exam Score")
```

Overall, the graph suggests that taking high school sociology may have a positive effect on students' scores across all three exams -- evidence by the higher median scores for the "yes" group (who have taken high school sociology). Taken together with the statistical significance from the MANOVA model, we can infer that hssoc dies indeed have an effect on the response variable (this does make intuitive sense given previous knowledge of a subject *should* improve future test performance).
