---
title: "STAC32 - Assignment 5"
author: "Michael Cowan - 997887032"
format: 
  html:
    embed-resources: true
    df-print: paged
editor: visual
---

```{r}
library(tidyverse)
library(devtools)
install_github("nxskok/smmr")
library(smmr)
```

# **1. Concrete**

**a)** Let's read in the data with read_csv.

```{r}
concrete_data <- read_csv("http://ritsokiguess.site/datafiles/ex14.23.csv")
concrete_data
```

**b)** Both a histogram and a QQ plot are appropriate here for our purposes (although the QQ plot will give us a more precise assessment of normality).

```{r}
# Histogram
ggplot(concrete_data, aes(x = strength)) +
  geom_histogram(bins = 6) +
  labs(x = "Strength (psi)", 
       y = "Frequency")
```

Having a look at the histogram, it appears that the data are right-skewed; we have a high 700 strength (psi) value at the high-tail of the distribution, and a cluster of values on the left side of the distribution. We can confirm the lack of normality (below) with a QQ plot -- which will give us a better look at the normality of the data than guesstimating with a histogram.

```{r}
# QQ plot
ggplot(concrete_data, aes(sample = strength)) +
   stat_qq(color = "blue") + stat_qq_line(color = "red")
```

As expected, we can see the points deviating significantly from the red line (the normal distribution). This will be problematic for doing t-tests, which operate on the assumption of normality.

**c)** We'll do sign-test here to check if the population median is a lower psi than 450 (we'll get a definitive answer from the p-value for the lower alternative sign-test).

```{r}
sign_test(concrete_data, strength, 450)
```

The p-value (p = 0.068) does not meet the 95% CI threshold for significance. Thus, based on this sign test, we cannot confidently say that there is evidence of a problem (though the value is subjectively quite close to 0.05). The median 7-day flexural strength cannot be confirmed to be lower than 450 psi, thus we fail to reject the null hypothesis (that the median flexural strength = 450 psi).

**d)** We'll compute a 99% CI with the smmr package functions.

```{r}
# 99% confidence interval for the median
ci_median(concrete_data, strength, conf.level=0.99)
```

Notably, the 99% confidence interval for the median 7-day flexural strength of the concrete (374.0084 - 459.9969) does encompasses the 450 psi value, as well as values above 450 psi (albeit not very many, ranging from 450-459.9969). This would seem to be in line with the default sign-test (95% CI) I conducted previously, which had a p-value close to significance suggesting the median could be lower than 450.

**e)** I'll generate and plot the bootstrap sampling distribution for the sample mean of the concrete strength data.

```{r}
# Bootstrap Sampling Distribution
set.seed(997887032)
tibble(sim = 1:1000) %>% 
  rowwise() %>% 
  mutate(bootstrap_sample = list(sample(concrete_data$strength, replace = TRUE))) %>% 
  mutate(bootstrap_mean = mean(bootstrap_sample)) %>% 
  ggplot(aes(x=bootstrap_mean)) + geom_histogram(bins=15) -> concrete_g

# Display the histogram
concrete_g
```

Given that the bootstrap sampling distribution of the sample mean appears to be normal, the Central Limit Theorem would suggest that parametric methods might still be appropriate -- irrespective of the original distribution of the concrete strength data, which led us to conduct a sign-test. Ultimately, while the original data did not appear to be normal, a t-test may be appropriate for this data.

# **2. Blood viscosity**

**a)** I'll read in the data with read_csv.

```{r}
blood_data <- read_csv("http://ritsokiguess.site/datafiles/xmp14.10.csv")
blood_data
```

**b)** First,we'll make a histogram.

```{r}
ggplot(blood_data, aes(x = cholesterol)) + 
  geom_histogram(bins = 8) + 
  labs(x = "Cholesterol Level", y = "Frequency")
```

**c)** Now for a Normal Quantile Plot.

```{r}
ggplot(blood_data, aes(sample = cholesterol)) +
   stat_qq(color = "blue") +
   stat_qq_line(color = "red")

```

**d)** Non-normality shows up on a histogram when points are clustered around a value other than the central peak/there are long tails at either end of the distribution. In the case of the cholesterol data, there is evidence of a slight right-skew (the tail is visible) due to some higher cholesterol values -- apparent when setting the histogram bins = 8.

This is also reflected in the QQ plot. While most of the lower-middle points seem to fall on and around the reference line (which represents normality), the points in the tails have a more pronounced deviation from the line -- especially the right tail -- reflecting the same non-normality observed in the histogram. Specifically, the upper range of the cholesterol data are greater than what we might expected if the values adhered to a normal distribution.

Ultimately, it would seem that the histogram and the QQ plot both indicate that the higher cholesterol data points may be the primary cause of deviations from normality in the data.
